---
title: "MMA 865 Final Project"
output: html_document
date: "Summer 2017"
author: "My Team Name"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
test<- "test"
```


```{r libraries}
library(tidyverse)
library(data.table)
```


## Load the data

```{r}
in_path = '/global/project/queens-mma/scene-csv/sample03/clean/'

# Note: read_csv() is a the preferred way to read in a CSV file in the tidyverse.
# Howewever, fread is 10 times faster. And why should *I* wait?

scene_mbr_dim <-fread(paste(in_path, 'scene_mbr_dim.csv', sep=""), sep=",")
scene_mbr_acct_dim <-fread(paste(in_path, 'scene_mbr_acct_dim.csv', sep=""), sep=",")
scene_pt_fact <-fread(paste(in_path, 'scene_pt_fact.csv', sep=""), sep=",")

# Note: you probably won't need these tables, but in case you do, here they are:
# scene_pt_tp_dim <-read_csv(paste(in_path, 'scene_pt_tp_dim.csv', sep=""))
# iwd_time <-read_csv(paste(in_path, 'iwd_time.csv', sep=""))
```

Note: the above loads the "small" data set. To load the full dataset instead, change the `in_path` to say "full" instead of `sample03`.

View the first few rows of some of the tables:

```{r}
library(dplyr)

#group by and arrange in desc order so as to have the latest updated account entry
scene_dim <- scene_mbr_dim %>% 
  group_by(scene_mbr_key,scene_mbr_seq_num) %>%
  arrange(desc(scene_mbr_seq_num)) %>%
  ungroup()
#remove the seq number
scene_dim_rm <- select(scene_dim, -2)
#remove the duplicate entries
scene_dim_dis <- distinct(scene_dim_rm, scene_mbr_key, .keep_all= TRUE)

scene_dim_test <- select(scene_dim, -2) 

#inner join the account table and scene_dim_dis table 
scene_inner_join<-inner_join(scene_dim_dis,scene_mbr_acct_dim,by=c("scene_mbr_key"="prim_scene_mbr_key"))

#remove NAs from transactional table (using base R)
scene_fact1 <-scene_pt_fact %>% filter(!is.na(txn_amt))


#group transaction table by the month and year and summarize the transaction amount 
scene_fact2 <- scene_fact1 %>%
  group_by(scene_mbr_key,anul_clndr_code,mo_clndr_code)%>%
  summarise(txn_amt = sum(txn_amt))%>%
  ungroup()

#join scene_fact and other tables
scene_join <- left_join(scene_fact2,scene_inner_join,by=c("scene_mbr_key"="scene_mbr_key"))

```


## Analysis

Go forth with your analysis! Use `dplyr` to manipulate the data, `ggplot2` to plot/graph, `rpart` (or similar) for building classifiers, `cluster` for clustering, and all your other training. Good luck, we're all counting on you.